---
title: "MetaR2M Tutorial"
author: "Zhichao Xu, and Peng Wei"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{MetaR2M Tutorial}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

## Introduction 
The [MetaR2M](https://github.com/zhichaoxu04/MetaR2M) `R` package presents a groundbreaking framework for conducting meta-analyses in high-dimensional settings, specifically for evaluating the R2-based total mediation effect. This innovative tool supports both fixed-effects and random-effects models, accommodating various research designs and high-throughput technologies. It is uniquely designed to handle the complexities inherent in high-dimensional data, a common feature in contemporary biomedical research.

Meta-analysis is a statistical technique used to combine the results of multiple studies to arrive at a comprehensive understanding of a particular field or topic. This method is especially prevalent in fields like medicine, psychology, and social sciences, where individual studies might have varying outcomes or small sample sizes. By aggregating data from several studies, a meta-analysis can provide more robust conclusions, identify patterns, and offer insights that might not be apparent from individual studies.

Fixed-effects models require the assumption that the true effects of interest are identical across all studies or cohorts. Random-effects models are used when there is heterogeneity across the studies included in the meta-analysis. In meta-analysis, the choice between a fixed-effect and a random-effects model is fundamental and depends on the underlying assumptions about the nature of the effects being analyzed.

The fixed-effect model operates under the assumption that there is one true effect size that is common to all the studies being analyzed. In essence, it posits that any observed differences in effect sizes across studies are solely due to sampling error. This model is particularly appropriate when the meta-analysis includes studies that are highly similar in terms of participants, interventions, and outcomes, suggesting that they are all tapping into the same underlying effect. The key advantage of this model is its simplicity and increased statistical power due to the assumption of a single underlying effect. However, its major limitation is that it cannot account for variability beyond chance among different studies, which can lead to biased results if this assumption is violated.

On the other hand, the random-effects model acknowledges and accommodates heterogeneity in effect sizes across studies. This model assumes that the studies in the meta-analysis are estimating different, yet related, effect sizes. These differences could arise from variations in study populations, methodologies, or other contextual factors. The random-effects model includes both within-study sampling error and between-study variance in its calculations. This approach is more flexible and realistic in scenarios where study heterogeneity is expected. It provides a more generalized conclusion, applicable to a broader context beyond the specific studies included in the meta-analysis. The trade-off, however, is that this model often has less statistical power compared to the fixed-effect model due to the additional variance component that needs to be estimated.


## Get started
Download and install following required R packages:

- Download [MetaR2M](https://github.com/zhichaoxu04/MetaR2M) package from Github using:

<!-- -->

    git clone https://github.com/zhichaoxu04/crSKAT.git

- Or, install [MetaR2M](https://github.com/zhichaoxu04/MetaR2M) package in R directly

  - First, install [devtools](https://devtools.r-lib.org) in R from CRAN:
    ``` r
    install.packages("devtools")
    ```
  - Then, install [MetaR2M](https://github.com/zhichaoxu04/MetaR2M) using the `install_github` function and load the package:
    ``` r
    devtools::install_github("zhichaoxu04/MetaR2M")
    library(MetaR2M)
    ```
- Make sure that all the required packages have been installed or updated. Here are some of the required packages:
  - [RsqMed](https://cran.r-project.org/web/packages/RsqMed/index.html): An implementation of calculating the R-squared measure as a total mediation effect size measure and its confidence interval for moderate- or high-dimensional mediator models. It gives an option to filter out non-mediators using variable selection methods. The original R package is directly related to the paper [Yang et al (2021)](https://pubmed.ncbi.nlm.nih.gov/34425752/). The new version contains a choice of using cross-fitting, which is computationally faster. The details of the cross-fitting method are available in our paper [Xu et al (2023)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9934518/).
  - [SIS](https://cran.r-project.org/web/packages/SIS/index.html): Variable selection techniques are essential tools for model selection and estimation in high-dimensional statistical models. Through this publicly available package, they provide a unified environment to carry out variable selection using iterative sure independence screening (SIS) ([Fan and Lv (2008)](https://academic.oup.com/jrsssb/article/70/5/849/7109492)) and all of its variants in generalized linear models ([Fan and Song (2009)](https://projecteuclid.org/journals/annals-of-statistics/volume-38/issue-6/Sure-independence-screening-in-generalized-linear-models-with-NP-dimensionality/10.1214/10-AOS798.full)) and the Cox proportional hazards model ([Fan, Feng and Wu (2010)](https://projecteuclid.org/ebooks/institute-of-mathematical-statistics-collections/Borrowing-Strength--Theory-Powering-Applications--A-Festschrift-for/chapter/High-dimensional-variable-selection-for-Coxs-proportional-hazards-model/10.1214/10-IMSCOLL606)).
  - [HDMT](https://cran.r-project.org/web/packages/HDMT/index.html): A multiple-testing procedure for high-dimensional mediation hypotheses. Mediation analysis is of rising interest in epidemiology and clinical trials. Methods used in the package refer to [James Y. Dai, Janet L. Stanford & Michael LeBlanc (2020)](https://www.tandfonline.com/doi/full/10.1080/01621459.2020.1765785).
  - [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html): A Grammar of Data Manipulation: a fast, consistent tool for working with data frame like objects, both in memory and out of memory.
  - [meta](https://cran.r-project.org/web/packages/meta/index.html): User-friendly general package providing standard methods for meta-analysis and supporting Schwarzer, Carpenter, and Rücker, ["Meta-Analysis with R" (2015)](https://link.springer.com/book/10.1007/978-3-319-21416-0).

## Toy Example

Let's suppose we aim to perform a meta-analysis using data from four distinct cohorts within a substantial genetic database. Our goal is to assess the mediating impact of gene expression on the age-related variations observed in systolic blood pressure (BP). For illustrative purposes, we will utilize the example data preloaded in our package as a representative sample for this analysis.

The outcome Y will be the systolic BP generated from a normal distribution with mean 100 and standard error 15. The exposure X will be the age generate from a normal distribution with mean 60 with standard error 10.

```{r, , eval=FALSE}
# Install the package
devtools::install_github("zhichaoxu04/MetaR2M")
```

```{r example}
# Load the dataset in the package
library(MetaR2M)
library(dplyr)
data("exampleData")

# Check the data structure
# Column names
colnames(exampleData$study1)[1:5]
# Dimension
dim(exampleData$study1)
# Summary stats
summary(exampleData$study1$Y)
summary(exampleData$study1$X)
```

```{r, message=FALSE}
# Use the first study to run CF-OLS to get the R2M
Y <- exampleData$study1$Y
M <- exampleData$study1 %>% dplyr::select(dplyr::starts_with("M_"))
X<- exampleData$study1$X
result1 <- CF_OLS(Y=Y, M=M, X=X, FDR=FALSE)
result1$output
```

This result could be interpreted as there is no mediating effects between age and systolic BP through our simulated 50 mediators because the confidence interval includes 0. This makes sense because we simulated these variables from independent distributions. After repeating this procedure, we could get the etstimate of R2M and its standard error. Then we could proceed to the meta-analysis of the R2M.

```{r}


```



## Notes:
Questions or novel applications? Please let me know! Contact information can be found in the [Github](https://github.com/zhichaoxu04). 
